\subsection{Structured contracts}
\label{sec:struc}

The structured contract framework~\cite{structured} is a formalism for specifying and demonstrating
the integrity of the implementation of a stateful contract on $\LEDGER$.
We give the definition here for self-containment and in order to introduce the
appropriate notation.
A structured contract includes a small-steps semantics specification, as well as
a ledger representation of its state and input. The ledger representation
is a pair of functions: one which computes the contract state from
a given UTxO state (or fails), and another which computes the input to the contract
for a given transaction.

For a given valid $\LEDGER$ step, the representation functions must compute a
valid step in the structured contract specification given that the starting
UTxO state corresponds to a contract state. This integrity constraint is
expressed as a proof obligation for the instantiation of a structured contract.
This design guarantees that no invalid contract state updates are ever possible
on the ledger.

Suppose $\STRUC \subseteq (\emptytypeT \times \State \times \Input \times \State)$
is a small-step transition system.
Let $\pi : \UTxO \to \State \cup \emptytypeT$ and $\pi_{\Tx} : \Tx \to \Input$
be functions such that :
\begin{equation*}
\inference
{
    \\~\\
    \pi~u~ \neq~ \emptytype &
    {
    \begin{array}{c}
        e\\
    \end{array}
    }
    \vdash
    {
    \left(
        \begin{array}{r}
        \var{u} \\
        \end{array}
    \right)
    }
    \trans{ledger}{\var{t}}
    {
    \left(
        \begin{array}{r}
        \var{u'} \\
        \end{array}
    \right)
    }
    \\~\\
}
{
    (\pi~u'~ \neq~ \emptytype) ~~\wedge~~
    \begin{array}{r}
    \emptytype \\
    \end{array}
\vdash
    (\pi~u)
    \trans{struc}{\pi_{\Tx}~t}
    (\varUpdate{\pi~u'})
}
\end{equation*}

The triple $(\STRUC, \pi, \pi_{\Tx})$ is called a \emph{structured
contract}, and we denote it by $(\STRUC, \pi, \pi_{\Tx}) \succeq \LEDGER$.
Note that $\pi$ is function with an output that is a \emph{maybe} type, 
$\State \cup \emptytypeT$, where $\emptytypeT$ is a
singleton. When $\pi~u = \emptytype$, there is no
contract state corresponding to the ledger state $u$.
The block-level data is never exposed to user-defined scripts in this model,
so that the context of a structured contract is necessarily $\emptytype \in \emptytypeT$.
